Generative AI for Image Reconstruction with Intensity Interferometry: A First Attempt

### 2024-10-02 Phase_II/draft/ParaGan.tex
second edit

\section{The Parameter Analysis for GAN}
\begin{figure*}
	\centering
	\begin{subfigure}{\linewidth}
		\includegraphics[width=\linewidth]{fig/testing_image/image_0.png}
	\end{subfigure}
	\begin{subfigure}{\linewidth}
		\includegraphics[width=\linewidth]{fig/testing_image/image_16.png}
	\end{subfigure}
	\begin{subfigure}{\linewidth}
		\includegraphics[width=\linewidth]{fig/testing_image/image_35.png}
	\end{subfigure}
	\begin{subfigure}{\linewidth}
		\includegraphics[width=\linewidth]{fig/testing_image/image_38.png}
	\end{subfigure}
	\begin{subfigure}{\linewidth}
		\includegraphics[width=\linewidth]{fig/testing_image/image_42.png}
	\end{subfigure}
	\begin{subfigure}{\linewidth}
		\includegraphics[width=\linewidth]{fig/testing_image/image_47.png}
	\end{subfigure}
	\caption{This set of figures shows the result of GAN for II. The left panels are signals using six baselines, which work as input for GAN. The first middle panel is the real image, also called ground truth, and GAN tries to mimic it during the training. The second middle panel is the reconstructed image, also called the predicted image, with trained GAN, and the end panel is the difference between the ground truth and the predicted image.}
	\label{fig:GAN}
\end{figure*}
Here, we discuss the parameters for the GAN structure, which is trained to reconstruct the image of stellar objects using II. Due to the adversarial nature of GANs, where the generator and discriminator compete in a min-max game, careful tuning of key parameters ensures that both networks converge properly.

\subsection{Data preparation}
First, we simulate fast-rotating stars, which are modeled as oblate spheroids with different radii and oblateness between 0.5 and 1. Different viewed angles have also been considered, with the effect of gravity darkening being linear dependence. The traced ellipses arise when integrating over the hour angle of the source. For the hyperparameter tuning and comparison of different telescopes, the total hour angle is about 11.5 hours. The ellipses are plotted and converted into grayscale images, resized, and stored as pure arrays to simplify further analysis. 
\begin{figure*}
	\centering
	\begin{subfigure}{0.33\linewidth}
		\includegraphics[width=\linewidth]{fig/moments/mom0.png}
		\caption{The monopole.}
		\label{fig:mom1}
	\end{subfigure}\hfill
	\begin{subfigure}{0.33\linewidth}
		\includegraphics[width=\linewidth]{fig/moments/mom1.png}
		\caption{The centroids along x direction.}
		\label{fig:mom2}
	\end{subfigure}\hfill
	\begin{subfigure}{0.33\linewidth}
		\includegraphics[width=\linewidth]{fig/moments/mom2.png}
		\caption{The centroids along y direction.}
		\label{fig:mom3}
	\end{subfigure}\hfill
	\caption{The monopole, centroid along x and y direction of ground truth and predicted image.}
	\label{fig:cen}
\end{figure*}
\begin{figure*}
	\centering
	\begin{subfigure}{0.33\linewidth}
		\includegraphics[width=\linewidth]{fig/moments/mom3.png}
		\caption{The 2nd order moment $\mu_{11}$.}
		\label{fig:mom4}
	\end{subfigure}\hfill
	\begin{subfigure}{0.33\linewidth}
		\includegraphics[width=\linewidth]{fig/moments/mom4.png}
		\caption{The 2nd order moment $\mu_{20}$.}
		\label{fig:mom5}
	\end{subfigure}\hfill
	\begin{subfigure}{0.33\linewidth}
		\includegraphics[width=\linewidth]{fig/moments/mom5.png}
		\caption{The 2nd order moment $\mu_{02}$.}
		\label{fig:mom6}
	\end{subfigure}\hfill
	\caption{The second-order central moments provide information about the size and shape of stellar objects.}
	\label{fig:struc}
\end{figure*}
\begin{figure*}
	\centering
	\begin{subfigure}{0.50\linewidth}
		\includegraphics[width=\linewidth]{fig/moments/mom6.png}
		\caption{The 3rd order moment $\mu_{30}$.}
		\label{fig:mom7}
	\end{subfigure}\hfill
	\begin{subfigure}{0.50\linewidth}
		\includegraphics[width=\linewidth]{fig/moments/mom7.png}
		\caption{The 3rd order moment $\mu_{03}$.}
		\label{fig:mom8}
	\end{subfigure}\hfill
	\begin{subfigure}{0.50\linewidth}
		\includegraphics[width=\linewidth]{fig/moments/mom8.png}
		\caption{The 3rd order moment $\mu_{21}$.}
		\label{fig:mom9}
	\end{subfigure}\hfill
	\begin{subfigure}{0.50\linewidth}
		\includegraphics[width=\linewidth]{fig/moments/mom9.png}
		\caption{The 3rd order moment $\mu_{12}$.}
		\label{fig:mom10}
	\end{subfigure}\hfill
	\caption{This figure shows the 3rd-order central moments for reconstructed images and represents the brightness distribution.}
	\label{fig:moments}
\end{figure*}

First, Salt and Pepper noise is introduced at a rate of $\alpha$ (usually 0.5\%). Then, the images are resized, and the mean is subtracted. A two-dimensional Fast Fourier Transform, as well as a Fourier shift, is applied, which yields a complex number of every pixel. As II does not measure the phase, the absolute value is calculated (shown in fig.~ \ref{fig:ft} on a linear and logarithmic scale for visualization). The sparse sampling is also introduced, namely by a pixel-wise multiplication between the absolute valued Fourier transformed image and the sparse sampling map. The result is a map in the Fourier plane with several ellipses, also called the sparse sampling map. The fig.~\ref{fig:ft_base} shows the sparse sampling for fig.~\ref{fig:image} as the source with four telescopes (fig.~\ref{fig:teles}). Finally, the pixels are normalized and converted to 8-bit integers. This image represents the sparsely sampled complex visibility, as it can be measured with II. The image shown in figure \ref{fig:ft_base} is the input image for the GAN, which also requires the ground truth image. So, the simulated stars are also resized using the same algorithm and converted to 8-bit integers to reduce the bias. The GAN must know the ground truth corresponding to an input image. Hence, they are merged side-by-side, shown in figure \ref{fig:GANinput}, and used as input to train the GAN. This procedure is conducted for all the simulated stars, with 10\% acting as test data, 10\% as validation data, and the remaining 80\% as training data.

@@

The learning rate of the optimizer is a measure of how much the model changes per iteration. Small learning rates can lead to underfitting, and large learning rates can make the model unstable. It is therefore important to carefully choose the learning rate \citep{murphy2022probabilistic}. Figure \ref{fig:Plot_learning_rate_loss} shows the effect of the learning rate on both generator and discriminator loss for three different values of it. Both loss functions have fewer outliers with decreasing learning rates, which is expected, as there is less change in the models with low learning rates. Even though all models stabilize at the same level, low learning rates are preferred.

The kernel size is the size of the kernel in the convolutional layer. It determines the amount of pixels that are linearly combined into a new pixel. A large kernel size allows one to catch features that are a few pixels away, but it also means that unrelated features might be introduced. The kernel size does not have a big effect on the loss functions, as shown in figure \ref{fig:Plot_kernel_size_loss}. Smaller kernel sizes seem to have more outliers, which means that either generator or discriminator gains an advantage. Therefore, a kernel size of 5 is preferred.

@@

The batch size defines the number of images that pass the network simultaneously. It is observed that smaller batch sizes can lead to better generalization \citep{prince2023understanding}. Figure \ref{fig:Plot_batchsize_loss} shows two different batch sizes. Here, when two images are processed simultaneously, there are fewer outliers, but since it increases the training time heavily, a batch size of 1 is used.

While training GANs, one can give the discriminator an advantage by increasing the number of training steps for it before returning to the generator training. It can increase the performance of the model. However, it increases the training time as well and leads to a lower discriminator loss score, as shown in \ref{fig:Plot_discrep_loss}. At the same time, the generator score increases slightly, which should be no surprise. Because the generated images did not improve with higher discriminator training, they are usually trained in the same amount as the generator.

Finally, the amount of sparse sampling can be varied, giving the model access to more pixels. More telescopes automatically lead to more baselines and, hence, more pixels that are available. Figure \ref{fig:Plot_telescopes_loss} shows the loss functions for different numbers of telescopes. There is a huge disparity, which can partly be explained by the fact that the relation between telescopes and baselines is not linear. The Fourier plane can be sampled sixfold when using four telescopes in contrast to only two telescopes. In the latter case, both the generator and discriminator are not trained smoothly, as seen by the outliers. It is already better with three telescopes and very promising when using four telescopes. The degree of sparse sampling seems to have the biggest effect of all hyperparameters.

>>>>

\section{The Parameter Analysis for GAN}
\begin{figure*}
	\centering
	\begin{subfigure}{\linewidth}
		\includegraphics[width=\linewidth]{fig/testing_image/image_0.png}
	\end{subfigure}
	\begin{subfigure}{\linewidth}
		\includegraphics[width=\linewidth]{fig/testing_image/image_16.png}
	\end{subfigure}
	\begin{subfigure}{\linewidth}
		\includegraphics[width=\linewidth]{fig/testing_image/image_35.png}
	\end{subfigure}
	\begin{subfigure}{\linewidth}
		\includegraphics[width=\linewidth]{fig/testing_image/image_38.png}
	\end{subfigure}
	\begin{subfigure}{\linewidth}
		\includegraphics[width=\linewidth]{fig/testing_image/image_42.png}
	\end{subfigure}
	\begin{subfigure}{\linewidth}
		\includegraphics[width=\linewidth]{fig/testing_image/image_47.png}
	\end{subfigure}
	\caption{This set of figures shows the result of GAN for II. The left panels are signals using six baselines, which work as input for GAN. The first middle panel is the real image, also called ground truth, and GAN tries to mimic it during the training. The second middle panel is the reconstructed image, also called the predicted image, with trained GAN, and the end panel is the difference between the ground truth and the predicted image.}
	\label{fig:GAN}
\end{figure*}
Here, we discuss the parameters for the GAN structure, which is trained to reconstruct the image of stellar objects using II. Due to the adversarial nature of GAN, where the generator and discriminator compete in a min-max game, careful tuning of key parameters ensures that both networks converge properly.

\subsection{Data preparation}
First, we simulate fast-rotating stars, which are modeled as oblate spheroids with different radii and oblateness between 0.5 and 1. Different viewed angles have also been considered, with the effect of gravity darkening being linear dependence. The traced ellipses arise when integrating over the hour angle of the source. For the hyperparameter tuning and comparison of different telescopes, the total hour angle is about 11.5 hours. The ellipses are plotted and converted into grayscale images, resized, and stored as pure arrays to simplify further analysis. 
\begin{figure*}
	\centering
	\begin{subfigure}{0.33\linewidth}
		\includegraphics[width=\linewidth]{fig/moments/mom0.png}
		\caption{The monopole.}
		\label{fig:mom1}
	\end{subfigure}\hfill
	\begin{subfigure}{0.33\linewidth}
		\includegraphics[width=\linewidth]{fig/moments/mom1.png}
		\caption{The centroids along x direction.}
		\label{fig:mom2}
	\end{subfigure}\hfill
	\begin{subfigure}{0.33\linewidth}
		\includegraphics[width=\linewidth]{fig/moments/mom2.png}
		\caption{The centroids along y direction.}
		\label{fig:mom3}
	\end{subfigure}\hfill
	\caption{This set of figures shows the comparison of monopole, x-centroid, and y-centroid for ground truth and predicted images generated by trained GAN.}
	\label{fig:cen}
\end{figure*}
\begin{figure*}
	\centering
	\begin{subfigure}{0.33\linewidth}
		\includegraphics[width=\linewidth]{fig/moments/mom3.png}
		\caption{The 2nd order moment $\mu_{11}$.}
		\label{fig:mom4}
	\end{subfigure}\hfill
	\begin{subfigure}{0.33\linewidth}
		\includegraphics[width=\linewidth]{fig/moments/mom4.png}
		\caption{The 2nd order moment $\mu_{20}$.}
		\label{fig:mom5}
	\end{subfigure}\hfill
	\begin{subfigure}{0.33\linewidth}
		\includegraphics[width=\linewidth]{fig/moments/mom5.png}
		\caption{The 2nd order moment $\mu_{02}$.}
		\label{fig:mom6}
	\end{subfigure}\hfill
	\caption{The second-order central moments provide information about the size and shape of stellar objects. This set of figures shows all the second-order central moments for ground truth and predicted images generated by trained GAN.}
	\label{fig:struc}
\end{figure*}
\begin{figure*}
	\centering
	\begin{subfigure}{0.50\linewidth}
		\includegraphics[width=\linewidth]{fig/moments/mom6.png}
		\caption{The 3rd order moment $\mu_{30}$.}
		\label{fig:mom7}
	\end{subfigure}\hfill
	\begin{subfigure}{0.50\linewidth}
		\includegraphics[width=\linewidth]{fig/moments/mom7.png}
		\caption{The 3rd order moment $\mu_{03}$.}
		\label{fig:mom8}
	\end{subfigure}\hfill
	\begin{subfigure}{0.50\linewidth}
		\includegraphics[width=\linewidth]{fig/moments/mom8.png}
		\caption{The 3rd order moment $\mu_{21}$.}
		\label{fig:mom9}
	\end{subfigure}\hfill
	\begin{subfigure}{0.50\linewidth}
		\includegraphics[width=\linewidth]{fig/moments/mom9.png}
		\caption{The 3rd order moment $\mu_{12}$.}
		\label{fig:mom10}
	\end{subfigure}\hfill
	\caption{This set of figures shows all the 3rd-order central moments for ground truth and predicted images generated by trained GAN. It calculates the skewness and evaluates the brightness distribution of images.}
	\label{fig:moments}
\end{figure*}

First, Salt and Pepper noise is introduced at a rate of $\alpha$ (usually 0.5\%). Then, the images are resized, and the mean is subtracted. A two-dimensional Fast Fourier Transform, as well as a Fourier shift, is applied, which yields a complex number of every pixel. As II does not measure the phase, the absolute value is calculated (shown in Fig.~\ref{fig:ft} on a linear and logarithmic scale for visualization). The sparse sampling is also introduced, namely by a pixel-wise multiplication between the absolute valued Fourier transformed image and the sparse sampling map. The result is a map in the Fourier plane with several ellipses, also called the sparse sampling map. Fig.~\ref{fig:ft_base} shows the sparse sampling for Fig.~\ref{fig:image} as the source with four telescopes (Fig.~\ref{fig:teles}). Finally, the pixels are normalized and converted to 8-bit integers. This image represents the sparsely sampled complex visibility, as it can be measured with II. The image shown in Fig.~\ref{fig:ft_base} is the input image for the GAN, which also requires the ground truth image. So, the simulated stars are also resized using the same algorithm and converted to 8-bit integers to reduce the bias. The GAN must know the ground truth corresponding to an input image. Hence, they are merged side-by-side, shown in Fig.~\ref{fig:GANinput}, and used as input to train the GAN. This procedure is conducted for all the simulated stars, with 10\% acting as test data, 10\% as validation data, and the remaining 80\% as training data.

@@

The learning rate of the optimizer is a measure of how much the model changes per iteration. Small learning rates can lead to underfitting, and large learning rates can make the model unstable. It is therefore important to carefully choose the learning rate \citep{murphy2022probabilistic}. Fig.~\ref{fig:Plot_learning_rate_loss} shows the effect of the learning rate on both generator and discriminator loss for three different values of it. Both loss functions have fewer outliers with decreasing learning rates, which is expected, as there is less change in the models with low learning rates. Even though all models stabilize at the same level, low learning rates are preferred.

The kernel size is the size of the kernel in the convolutional layer. It determines the amount of pixels that are linearly combined into a new pixel. A large kernel size allows one to catch features that are a few pixels away, but it also means that unrelated features might be introduced. The kernel size does not have a big effect on the loss functions, as shown in Fig.~\ref{fig:Plot_kernel_size_loss}. Smaller kernel sizes seem to have more outliers, which means that either generator or discriminator gains an advantage. Therefore, a kernel size of 5 is preferred.

@@

The batch size defines the number of images that pass the network simultaneously. It is observed that smaller batch sizes can lead to better generalization \citep{prince2023understanding}. Fig.~\ref{fig:Plot_batchsize_loss} shows two different batch sizes. Here, when two images are processed simultaneously, there are fewer outliers, but since it increases the training time heavily, a batch size of 1 is used.

While training GAN, one can give the discriminator an advantage by increasing the number of training steps before returning to the generator training. It can increase the performance of the model. However, it increases the training time as well and leads to a lower discriminator loss score, as shown in Fig.~\ref{fig:Plot_discrep_loss}. At the same time, the generator score increases slightly, which should be no surprise. Because the generated images did not improve with higher discriminator training, they are usually trained in the same amount as the generator.

Finally, the amount of sparse sampling can be varied, giving the model access to more pixels. More telescopes automatically lead to more baselines and, hence, more pixels that are available. Fig.~\ref{fig:Plot_telescopes_loss} shows the loss functions for different numbers of telescopes. There is a huge disparity, which can partly be explained by the fact that the relation between telescopes and baselines is not linear. The Fourier plane can be sampled sixfold when using four telescopes in contrast to only two telescopes. In the latter case, both the generator and discriminator are not trained smoothly, as seen by the outliers. It is already better with three telescopes and very promising when using four telescopes. The degree of sparse sampling seems to have the biggest effect of all hyperparameters.

### 2024-10-02 Phase_II/draft/Result.tex
second edit

\section{The Reconstructed Image with GAN}
In this section, we will first discuss phase retrieval using hyperparameters, as mentioned earlier, followed by an analysis where multiple sources are trained simultaneously. The best performance for image reconstruction has been observed with a learning rate of $2 \cdot 10^{-4}$, a kernel size of 5x5, and equal noise percentages in original and generated images. The batch size selected is 1, and the discriminator-generator receives equal training.
\subsection{The Predicted Image with Trained GAN}
The success of the GAN in training the model for Intensity Interferometry to reconstruct images of fast-rotating stars is demonstrated in Fig.~\ref{fig:GAN}. The GAN was trained on training datasets for 60,000 steps. After training, the GAN was tested on different validation datasets, producing predicted images of a fast-rotating star. Fig.~\ref{fig:GAN} presents a set of four combined images demonstrating the GAN's performance in reconstructing the shape, size, and brightness distribution of the fast-rotating star using II. The left panel shows the signals covered by six baselines, which serve as the input for the generator to train the GAN. The first middle panel displays the real image, or ground truth, which the discriminator uses to differentiate between real and generated images from the generator. During training, the GAN aims to mimic this ground truth. The second middle panel depicts the reconstructed image, or predicted image, produced by the trained GAN. This panel illustrates the success of the GAN in image reconstruction using II. The right panel shows the difference between the ground truth and the predicted image. This difference should be minimized for high-precision image reconstruction with the application of GAN on II. The predicted images of Fig.~\ref{fig:GAN} showed positive results, accurately providing visual information about the source's size and shape, as well as the distributed brightness. This image reconstruction was achieved using only six baselines. However, the results can be further improved by increasing the number of telescopes to cover more (u, v) planes, making the existing and upcoming Cherenkov Telescope Array Observatory (CTAO) an ideal candidate in collaboration with existing CTA. 
\subsection{Evaluation of GAN using moments}
After the success of GAN in reconstructing the image using II, there is a need for evaluation of it. So, we use the image moments. Image moments are statistical properties that provide information about the reconstructed shape, size, and intensity distribution of the objects in the image. These moments help quantify key features like the position, orientation, and distribution of brightness in the image, allowing for an objective evaluation of how well the predicted image corresponds to the actual object. We will analyze the consistency and accuracy of the GAN-generated image by comparing its moments to those of the ground truth. It provides a reliable framework for evaluating reconstruction quality, as image moments highlight the subtle differences in geometric and intensity properties that might not be visually apparent.

@@

The zeroth order raw moment, which is the total intensity of an image and is called the monopole. It sums up all the pixel values across the image, providing an overall intensity value. So, the study of monopole provides the total flux of fast-rotating stars here. According to eqn.~\ref{eqn:Mom}, the monopole of a image is calculated as 
\begin{equation}
	M_{00} = \sum_{x} \sum_{y} I(x, y).
\end{equation}
Fig.~\ref{fig:mom1} shows the monopole for 50 different reconstructed images. It shows the linear behavior for the monopole between the ground truth (the real image) along the x-axis and the predicted image (reconstructed image) along the y-axis, which is obvious for different shape-size sources. This result explains the similarity between the total intensity of both images. It ensures that the predicted image has the approximately correct total brightness (the flux) compared to the ground truth. However, monopole does not explain the shape, size, and brightness distribution of fast-rotating stars. So, there is a need for higher-order moments.

The center of mass for the fast-rotating star or any other stellar object is calculated using the centroid (in x and y directions). These centroids are given in terms of first-order raw moment and monopole as
\begin{equation}
	\begin{aligned}
		&m_x = \frac{\sum_{x} x I(x,y)}{\sum_{x} \sum_{y} I(x, y)} = \frac{M_{10}}{M_{00}} \\
		&m_y = \frac{\sum_{y} y I(x,y)}{\sum_{x} \sum_{y} I(x, y)} = \frac{M_{01}}{M_{00}}
	\end{aligned}  
\end{equation}
Fig.~\ref{fig:mom2} and Fig.~\ref{fig:mom3} show the comparison of centroid along the x and y axis for 50 predicted images with respect to ground truths. The clustering of centroids in a given scale range for all the results explains that the reconstructed image correctly represents the spatial location of the fast-rotating star in compare with the ground truth.

@@

The information on brightness distribution is gathered using the skewness of the image by calculating third-order central moment ($\mu_{30}, \mu_{03}, \mu_{21}, \mu_{12}$) of images. Fig.~\ref{fig:moments} shows all third-order moments for the ground truth and reconstructed image. The skewness along the x and y axis ($\mu_{30}, \mu_{03}$) to test the GAN for II are acceptable, which can be seen in Fig.~\ref{fig:mom7} and Fig.~\ref{fig:mom8}, where a linear relation exists between ground truth and predicted image. However, the remaining higher moments $(\mu_{21}, \mu_{12})$ shown in Fig.~\ref{fig:mom9} and Fig.~\ref{fig:mom10} are not in good terms specially $\mu_{12}$.

\subsection{The reconstructed Parameters for object}
The centroids $(m_x, m_y)$ represents the center of the fast-rotating star. However, the calculated second-order central moment defines the orientation, semi-major axis, and eccentricity of the source \citep{teague1980image}. These parameters fully describe the ellipse that fits the image data based on the moments.

The orientation along the line of sight is defined as
\begin{equation}
	\theta = \frac{1}{2}\arctan \big(\frac{2\mu_{11}}{\mu_{20} - \mu_{02}}\big).
	\label{eqn:orn}
\end{equation}
The semi-major and semi-minor axis will be calculated according to
\begin{equation}
	\begin{aligned}
		&a = 2\sqrt{mp + \delta} \\
		&b = 2\sqrt{mp - \delta}
	\end{aligned}
	\label{eqn:semi}
\end{equation}
where,
\begin{equation}
	mp = \frac{\mu_{20} + \mu_{02}}{2}
	\label{eqn:mp}
\end{equation}
and
\begin{equation}
	\delta = \frac{\sqrt{4\mu_{11}^2 + (\mu_{20} - \mu_{02})^2}}{2}.	
	\label{eqn:delta}
\end{equation}
So, the eccentricity of the fast-rotating star in terms of axis value is 
\begin{equation}
	e = \sqrt{1 - a/b}.
	\label{eqn:eccen}
\end{equation}
Fig.~\ref{fig:recons} shows all zeroth to third order moments only for one ground and predicted image. In this figure, there is a red curve for both ground truth and predicted image, which is the representation of all second-order moments in terms of eqn.~\ref{eqn:orn} to eqn.~\ref{eqn:eccen} as orientation, axis value, and eccentricity. There is a cyan dot point as well, which represents the zeroth order moment for both images. There is also a third-order moment, which is expressed in terms of arrows in magenta and blue color to view the skewness along the x and y axes.

>>>>

\section{The Reconstructed Image with GAN}
In this section, we will first discuss phase retrieval using hyperparameters, as mentioned earlier, followed by an analysis where multiple sources are trained simultaneously. The best performance for image reconstruction has been observed with a learning rate of $2 \cdot 10^{-4}$, a kernel size of 5x5, and equal noise percentages in original and generated images. The batch size selected is 1, and the discriminator-generator receives equal training.
\subsection{The Predicted Image with Trained GAN}
The success of the GAN in training the model for Intensity Interferometry to reconstruct images of fast-rotating stars is demonstrated in Fig.~\ref{fig:GAN}. The GAN was trained on training datasets for 60,000 steps. After training, the GAN was tested on different validation datasets, producing predicted images of a fast-rotating star. Fig.~\ref{fig:GAN} presents a set of four combined images demonstrating the GAN's performance in reconstructing the shape, size, and brightness distribution of the fast-rotating star using II. The left panel shows the signals covered by six baselines, which serve as the input for the generator to train the GAN. The first middle panel displays the real image, or ground truth, which the discriminator uses to differentiate between real and generated images from the generator. During training, the GAN aims to mimic these ground truth images. The second middle panel depicts the reconstructed image, or predicted image, produced by the trained GAN. This panel illustrates the success of the GAN in image reconstruction using II. The right panel shows the difference between the ground truth and the predicted image. This difference should be minimized for high-precision image reconstruction with the application of GAN on II. The predicted images of Fig.~\ref{fig:GAN} showed positive results, accurately providing visual information about the source's size and shape, as well as the distributed brightness. This image reconstruction was achieved using only six baselines. However, the results can be further improved by increasing the number of telescopes to cover more (u, v) planes, making the existing and upcoming Cherenkov Telescope Array Observatory (CTAO) an ideal candidate in collaboration with existing CTA. 
\subsection{Evaluation of GAN using moments}
The reconstructed images are good enough visually to claim the success of GAN in reconstructing the image using II. However, there is a need for statistical evaluation of this predicted image to guarantee success. So, we use image moments as a statistical method. Image moments are statistical properties that provide information about the reconstructed shape, size, and intensity distribution of the objects in the image. These moments help quantify key features like the position, orientation, and distribution of brightness in the image, allowing for an objective evaluation of how well the predicted image corresponds to the actual object. We will analyze the consistency and accuracy of the GAN-generated image by comparing its moments to those of the ground truth. It provides a reliable framework for evaluating reconstruction quality, as image moments highlight the subtle differences in geometric and intensity properties that might not be visually apparent.

@@

The zeroth order raw moment called the monopole is the total intensity of an image. It sums up all the pixel values across the image, providing an overall intensity value. So, the study of monopole provides the total flux of fast-rotating stars here. According to eqn.~\ref{eqn:Mom}, the monopole of an image is calculated as 
\begin{equation}
	M_{00} = \sum_{x} \sum_{y} I(x, y).
\end{equation}
Fig.~\ref{fig:mom1} shows the monopole for 50 different reconstructed images. It shows the linear behavior for the monopole between the ground truth (the real image) along the x-axis and the predicted image (reconstructed image) along the y-axis, which is obvious for different shape-size sources. This result explains the similarity between the total intensity of both images. It ensures that the predicted image has the approximately correct total brightness (the flux) compared to the ground truth. However, monopole does not explain the position, shape, size, and brightness distribution of fast-rotating stars; there is a need for higher-order moments.

The center of mass for the fast-rotating star or any other stellar object is calculated using the centroid (x-centroid and y-centroid). It represents the spatial position of the image and is calculated using first-order raw moment and monopole. The formulation of centroid along the x and y directions is
\begin{equation}
	\begin{aligned}
		&m_x = \frac{\sum_{x} x I(x,y)}{\sum_{x} \sum_{y} I(x, y)} = \frac{M_{10}}{M_{00}} \\
		&m_y = \frac{\sum_{y} y I(x,y)}{\sum_{x} \sum_{y} I(x, y)} = \frac{M_{01}}{M_{00}}
	\end{aligned}  
\end{equation}
Fig.~\ref{fig:mom2} and Fig.~\ref{fig:mom3} show the comparison of the x-centroid and y-centroid for 50 predicted images with respect to ground truths, respectively. The clustering of centroids in a given scale range for all the results explains that the reconstructed image correctly represents the spatial location of the fast-rotating star in compare with the ground truth.

@@

The information on brightness distribution is gathered using the skewness of the image by calculating third-order central moment ($\mu_{30}, \mu_{03}, \mu_{21}, \mu_{12}$) of images. Fig.~\ref{fig:moments} shows all third-order moments for the ground truth and reconstructed image. The skewness along the x and y axis ($\mu_{30}, \mu_{03}$) to test the GAN for II are acceptable, which can be seen in Fig.~\ref{fig:mom7} and Fig.~\ref{fig:mom8}, where a linear relation exists between ground truth and predicted image. However, the remaining higher moments $(\mu_{21}, \mu_{12})$ shown in Fig.~\ref{fig:mom9} and Fig.~\ref{fig:mom10} are not in good terms specially $\mu_{12}$. Further improvement is possible and needs to be studied.

\subsection{The reconstructed Parameters for object}
The centroids $(m_x, m_y)$ represent the center of the fast-rotating star and spatial location on the image. However, the calculated second-order central moment defines the orientation, semi-major axis, and eccentricity of the source \citep{teague1980image}. These parameters fully describe the ellipse that fits the image data based on the moments.

The orientation of a fast-rotating star along the line of sight is defined in terms of second-order central moments as
\begin{equation}
	\theta = \frac{1}{2}\arctan \big(\frac{2\mu_{11}}{\mu_{20} - \mu_{02}}\big).
	\label{eqn:orn}
\end{equation}
The semi-major and semi-minor axis of the stellar object will be calculated with the help of second-order central moment again and denoted as a and b here, respectively
\begin{equation}
	\begin{aligned}
		&a = 2\sqrt{mp + \delta} \\
		&b = 2\sqrt{mp - \delta}
	\end{aligned}
	\label{eqn:semi}
\end{equation}
where,
\begin{equation}
	mp = \frac{\mu_{20} + \mu_{02}}{2}
	\label{eqn:mp}
\end{equation}
and
\begin{equation}
	\delta = \frac{\sqrt{4\mu_{11}^2 + (\mu_{20} - \mu_{02})^2}}{2}.	
	\label{eqn:delta}
\end{equation}
With the help of the axis value, the eccentricity of the fast-rotating star is calculated as
\begin{equation}
	e = \sqrt{1 - a/b}.
	\label{eqn:eccen}
\end{equation}
Fig.~\ref{fig:recons} shows all zeroth to third order moments only for one ground and predicted image. In this figure, there is a red curve for both ground truth and predicted image, which is the representation of all second-order moments in terms of Eqn.~\ref{eqn:orn} to Eqn.~\ref{eqn:eccen} as orientation, axis value, and eccentricity. There is a cyan dot point as well, which represents the zeroth order moment for both images. The third-order moment is expressed in terms of arrows in magenta and blue color to view the skewness along the x and y axes.

### 2024-10-02 Phase_II/draft/conclusion.tex
second edit

\section{Conclusion}
The challenge of phase retrieval in Intensity Interferometry (II) has been effectively addressed through the application of machine learning techniques, particularly with Generative Adversarial Networks (GANs). Our study demonstrates that using a GAN on II data successfully recovers the size, shape, and brightness distribution of a fast-rotating star. The evaluation based on image moments, specifically the monopole, second, and third-order moments, supports the effectiveness of the GAN in achieving accurate image reconstruction from a one-night simulation of II using six baselines.

A critical factor in the performance of the reconstruction process is the extent of Fourier plane coverage, which is determined by the number of available telescopes and the total observing time. The brightness distribution can likely be reconstructed with even higher precision with a full night of observation using four telescopes. Future work could explore different observatory layouts, such as the Southern Cherenkov Telescope Array (CTA) layout, to evaluate the impact on image reconstruction quality.

This study has not considered the detector efficiencies, which could influence the signal-to-noise ratio (SNR) in practical scenarios. Incorporating these efficiencies in future work will be important for providing a more accurate estimate of the SNR and the robustness of the GAN in real-world applications. While conditional GANs have proven effective in this study, there are other methods for generating images from image data. Exploring and comparing alternative methods, as well as adapting the GAN architecture itself, could offer improvements. For instance, different loss functions could be implemented and compared to enhance reconstruction quality. Further testing is required to refine the GAN and make it more robust and reliable. However, our findings suggest that machine learning is a promising approach for phase reconstruction in II.

In summary, while the current implementation shows strong potential, further exploration and refinement of machine learning models are needed to optimize their application in phase retrieval and image reconstruction in Intensity Interferometry.

>>>>

\section{Conclusion}
Intensity Interferometry (II), a technique, is emerging after a long gap to overcome the challenges of Amplitude Interferometry on optical wavelength. It is rejuvenated with support from advanced facilities at the Imaging Cherenkov Telescope Arrays (ICTAs). These telescopes, equipped with large aperture areas and photon detectors capable of resolving signals within a nanosecond, are instrumental in capturing high-resolution images \citep{dravins2013optical}. The Major Atmospheric Gamma Imaging Cherenkov Telescope (MAGIC), one of the CTA in the world, is now also being used for II, observing stellar objects with its 17-meter diameter mirror \citep{lorenz2004status}. The telescope has already demonstrated the potential of II by measuring and comparing the diameters of individual stars \citep{abe2024performance}. Other existing arrays (Very Energetic Radiation Imaging Telescope Array System (VERITAS) and High Energy Stereoscopic System (HESS)) are also making progress as II when gamma-ray observation stops \citep{kieda2021veritas, zmija2022optical}. Future observations, made possible by more advanced, sensitive instrumentation and upcoming CTAO, promise to push the boundaries of what we can observe with optical range. However, there is a drawback of II in terms of phase loss of signal, capturing only its magnitude via photon correlation. Here, the challenge of phase retrieval in Intensity Interferometry (II) has been effectively addressed through the application of machine learning techniques, particularly with conditional Generative Adversarial Networks (cGAN). Our study demonstrates that the application of cGAN on II data successfully recovers the size, shape, and brightness distribution of a fast-rotating star. The evaluation based on image moments, specifically the monopole, second, and third-order moments, supports the effectiveness of the cGAN in achieving accurate image reconstruction from a one-night simulation of II using six baselines. A critical factor in the performance of the reconstruction process is the extent of Fourier plane coverage, which is determined by the number of available telescopes and the total observing time. The brightness distribution can likely be reconstructed with even higher precision with a full night of observation using four telescopes. Future work could explore different observatory layouts, such as the Southern Cherenkov Telescope Array (CTA) layout, to evaluate the impact on image reconstruction quality.

While the results of this study demonstrate the significant potential of machine learning, particularly cGAN, for image reconstruction in II, several aspects need further refinement. Detector efficiencies, which impact the signal-to-noise ratio (SNR) in real-world scenarios, have yet to be incorporated, and addressing them will be crucial for more accurate SNR estimation. Incorporating these efficiencies in future work will be important for providing a more accurate estimate of the SNR and the robustness of the GAN in real-world applications. While cGAN has proven effective in this study, there are other methods for generating images from image data. Exploring and comparing alternative methods, as well as adapting the GAN architecture itself, could offer improvements. For instance, different loss functions could be implemented and compared to enhance reconstruction quality. Further testing is required to refine the GAN and make it more robust and reliable. However, our findings suggest that machine learning is a promising approach for phase reconstruction in II.

In summary, though cGAN has shown promising results, further testing and improvements are necessary to make the model more robust and reliable in practical applications. Overall, this study highlights the potential of machine learning approaches for advancing phase retrieval and high-resolution image reconstruction in II, suggesting exciting future directions in the application of AI to optical astronomy.
